---
title: "MS_Florencia"
author: "Pedro Jordano"
date: "6-19/3/2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE, cache=T)

```

## Datasets

The data are frequencies of _Prosopis_ tree-frugivore interactions at the individual tree level in Argentina.

```{r input, echo= TRUE}
library(ggplot2)
library(statnet)
library(igraph)
library(sna)
library(ergm)
library(ggnet)
library(GGally)
# source("multiplot.R")
main<- read.csv("../data/interaction_matrix.csv", header= T, sep= ",",
                dec = ".", na.strings = "NA")
str(main)
attr<- read.csv("../data/ecological_attributes.csv", header= T, sep= ",",
                dec = ".", na.strings = "NA")
str(attr)

```

## Some edits

Fixing the adjacency matrices from the input data. Subset the two sites (`Grazed` and `Ungrazed`) and extract the adjacency matrices for each site.

This extracts just the interaction data for total effect (Total effect interaction data is in cols 32-43). The covariates are extracted later.

```{r adj_mat, echo=T}
grazed<- main %>%  
          dplyr::filter(site=="grazed") %>% 
          select(1,32:43) 

ungrazed<- main %>%  
          dplyr::filter(site=="ungrazed") %>% 
          select(1,32:43) 

head(grazed); head(ungrazed)

```

These are labels for plotting, etc. To be checked by Florencia. 

```{r labels, echo=TRUE}
# Labels for family:
famlab<- c()
#
# Labels for subfamilies:
sfamlab<- c()
#
# Labels for feeding modes:
feedlab<- c()
#

```

## Plotting the interaction networks by site

For each site, with node labels, weighted adjacency matrices.

```{r plots_by_site, fig.width=9, eval=T}
# Plotting bipartite networks from adjacency matrix of two-mode network.
# Using ggplot2 -----------------------------------------------------------
# 
# Code for plotting a bipartite network in ggplot2
require(network)
require(ggplot2)
require(sna)
require(ggnet)

# Bipartite network initialization, starting from an adjacency matrix.
# Matrix from dataframe.
wG<-t(grazed[,2:13])
colnames(wG)<- grazed[,1]
netwG<-network::network(wG, matrix.type= 'bipartite', ignore.eval= T)

wUG<-t(ungrazed[,2:13])
colnames(wUG)<- ungrazed[,1]
netwUG<-network::network(wUG, matrix.type= 'bipartite', ignore.eval= T)

# Function to compute edge weights, scaled.
edge.weights<- function(M, x = 10) {
    # Transpose.
    M <- t(M)
    # Edge list and weights.
    M <- cbind(expand.grid(dimnames(M))[2:1], as.vector(M))
    # Discard null weights.
    M <- subset(M, M[, 3] != 0) 
    # Scaled weights.
    M.scaled <- x*log(M[, 3] + 1) / max(log(M[, 3] + 1))
    # Vector of edge weights.
    return(M.scaled)
}

# Pass the network, edge weights and mode to ggnet.
# Old:
# net = bipartite.network(as.matrix(net80), modes = c("Herb", "Trees"))
# Weighted bipartite networks
# 
# GRAZED
bipwG= network(wG,
              matrix.type = "bipartite",
              ignore.eval = FALSE,
              names.eval = "weights",
              modes = c("Frug", "Trees"))
# set colors for each mode
# col = c("actor" = "grey", "event" = "gold")

ggnet2(bipwG, node.size = 5,
       edge.size= edge.weights(wG, 5), edge.alpha= .25,
       label= TRUE, label.size= 3,
       color= "mode", palette = "Set2",
       shape= "mode")

# UNGRAZED
bipwUG= network(wUG,
              matrix.type = "bipartite",
              ignore.eval = FALSE,
              names.eval = "weights",
              modes = c("Frug", "Trees"))
# set colors for each mode
# col = c("actor" = "grey", "event" = "gold")

ggnet2(bipwUG, node.size = 5,
       edge.size= edge.weights(wUG, 5), edge.alpha= .25,
       label= TRUE, label.size= 3,
       color= "mode", palette = "Set2",
       shape= "mode")

```



#### Analyses

The Bernouilli or random model assumes that, for a given pair of nodes, the presence or absence of an edge between that pair is independent of the status of possible edges between any other pairs of nodes.

These three statistics can be used in `ergm` by specifying terms `altkstar`, `gwdegree`, or `gwesp`, respectively, in the model. 

For all these analyses I've followed:
Kolaczyk, E.D. and  Csárdi, G. 2014. _Statistical Analysis of Network Data with R_. Springer, Berlin. (ISBN 978-1-4939-0982-7). See `library(sand)`.

##### Build data

```{r data.build, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# Add the node attributes
# 
net80.g<- graph_from_incidence_matrix(m80, weighted= T, add.names=NULL)
n80.edg<- as_edgelist(net80.g)

# Subsetting the covariates from the main dataset for each year separately.
# 
m80_attr<-main %>%  
          dplyr::filter(year==1980) %>% 
          select(c(2,28:32)) 

# Create the attributes matrix
V(net80.g)$tree<-       m80_attr$treeidpauci
V(net80.g)$tree_mass<-  m80_attr$tree_mass
V(net80.g)$leaf_P<-     m80_attr$leaf_P
V(net80.g)$height<-     m80_attr$height
V(net80.g)$leaf_N<-     m80_attr$leaf_N
V(net80.g)$NP_ratio<-   m80_attr$NP_ratio

```


##### Correations among covariates

```{r correlations, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8}
# Correlations among tree cahracteristics
# 
mat<-m80_attr[,c(2:6)]
mm<- stats::cor(mat, use="pairwise.complete.obs")

# library(GGally)
p1<- ggcorr(data = NULL, cor_matrix = cor(mm[, -1], use = "everything"))

p2<- ggcorr(m80_attr[,c(2:6)], geom = "blank", label = TRUE, hjust = 0.75) +
        geom_point(size = 10, aes(color = coefficient > 0, 
                   alpha = abs(coefficient) > 0.5)) +
                   scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
                   guides(color = FALSE, alpha = FALSE)
multiplot (p1, p2, cols=2)

```


##### Set matrices and graphs for model testing

For now just the 1980 graph is analyzed.

```{r Bernouilli, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# Loading vertex attributes for 1980
# 
AA <- get.adjacency(net80.g)
v.attrs <- get.data.frame(net80.g, what="vertices")[1:42,]

library(ergm) # Will load package ’network’ as well.
m80.s <- network::as.network(as.matrix(AA),
                             directed=FALSE, bipartite=42) 
network::set.vertex.attribute(m80.s, "tree_mass",
                                v.attrs$tree_mass) 
network::set.vertex.attribute(m80.s, "leaf_P",
                                v.attrs$leaf_P) 
network::set.vertex.attribute(m80.s, "height",
                                v.attrs$height)
network::set.vertex.attribute(m80.s, "leaf_N",
                                v.attrs$leaf_N)
network::set.vertex.attribute(m80.s, "NP_ratio",
                                v.attrs$NP_ratio)

# Random Bernouilli model
my.ergm.bern <- formula(m80.s ~ edges)
my.ergm.bern
m80.s ~ edges
summary.statistics(my.ergm.bern)

# Models with no explanatory variables
# 
m80.ergm00 <- formula(m80.s ~ edges)                      # Bernoulli

m80.ergm01 <- formula(m80.s ~ edges + kstar(2)            # kstar
                            + kstar(3) + triangle)

m80.ergm02 <- formula(m80.s ~ edges     # Geometrically weighted degree count
                    + gwesp(1, fixed=TRUE))

summary.statistics(m80.ergm00)
summary.statistics(m80.ergm01)
summary.statistics(m80.ergm02)

```


```{r expl.variables, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# Models with explanatory variables. 1980.
#
m80.ergm1 <- formula(m80.s ~ edges + gwesp(log(3), fixed=TRUE)
                        + nodemain("tree_mass")
                        + nodemain("leaf_P")
                        + nodemain("height")
                        + nodemain("leaf_N")
                        + nodemain("NP_ratio")) 

m80.ergm2 <- formula(m80.s ~ edges + gwesp(log(3), fixed=TRUE)
                        + nodemain("tree_mass")
                #       + nodemain("leaf_P")
                        + nodemain("height"))
                #       + nodemain("leaf_N")
                #       + nodemain("NP_ratio")) 
m80.ergm3 <- formula(m80.s ~ edges + gwesp(log(3), fixed=TRUE)
                #       + nodemain("tree_mass")
                        + nodemain("leaf_P")
                #       + nodemain("height")
                        + nodemain("leaf_N")
                        + nodemain("NP_ratio")) 

m80.ergm11 <- formula(m80.s ~ edges + gwesp(log(3), fixed=TRUE)
                        + nodemain("tree_mass"))

m80.ergm12 <- formula(m80.s ~ edges + gwesp(log(3), fixed=TRUE)
                        + nodemain("leaf_N"))

```


```{r test.expl.variables, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# Model tests
# 
set.seed(42)
m80.ergm.fit00 <- ergm(m80.ergm00)
m80.ergm.fit02 <- ergm(m80.ergm02)
m80.ergm.fit1 <- ergm(m80.ergm1)    # Full model
m80.ergm.fit2 <- ergm(m80.ergm2)    # Just tree size
m80.ergm.fit3 <- ergm(m80.ergm3)    # Just leaf nutrient quality
m80.ergm.fit11 <- ergm(m80.ergm11)
m80.ergm.fit12 <- ergm(m80.ergm12)

# NOT RUN
# mcmc.diagnostics(m80.ergm.fit0)
# mcmc.diagnostics(m80.ergm.fit1)
# mcmc.diagnostics(m80.ergm.fit2)
# mcmc.diagnostics(m80.ergm.fit3)
# mcmc.diagnostics(m80.ergm.fit11)
# mcmc.diagnostics(m80.ergm.fit12)

anova.ergm(m80.ergm.fit00, test="Chi")
anova.ergm(m80.ergm.fit02, test="Chi")
anova.ergm(m80.ergm.fit1, test="Chi")
anova.ergm(m80.ergm.fit2, test="Chi")
anova.ergm(m80.ergm.fit3, test="Chi")
anova.ergm(m80.ergm.fit11, test="Chi")
anova.ergm(m80.ergm.fit12, test="Chi")

anova.ergm(m80.ergm.fit00, m80.ergm.fit1, test="Chi")
anova.ergm(m80.ergm.fit02, m80.ergm.fit1, test="Chi")
anova.ergm(m80.ergm.fit1, m80.ergm.fit2, test="Chi")
anova.ergm(m80.ergm.fit1, m80.ergm.fit3, test="Chi")
anova.ergm(m80.ergm.fit2, m80.ergm.fit3, test="Chi")

summary.ergm(m80.ergm.fit00)
summary.ergm(m80.ergm.fit02)
summary.ergm(m80.ergm.fit1)
summary.ergm(m80.ergm.fit2)
summary.ergm(m80.ergm.fit3)
summary.ergm(m80.ergm.fit11)
summary.ergm(m80.ergm.fit12)

```


```{r AICw_ExplVars}

## Compute table
require(MuMIn)
tt<- round(BIC(m80.ergm.fit00, m80.ergm.fit02, m80.ergm.fit1, m80.ergm.fit2,
               m80.ergm.fit3, m80.ergm.fit11, m80.ergm.fit12), 10)
ttt<- round(Weights(AICc(m80.ergm.fit00, m80.ergm.fit02, m80.ergm.fit1,
                         m80.ergm.fit2, m80.ergm.fit3, m80.ergm.fit11,
                         m80.ergm.fit12)), 10)
model.summ<- arrange(data.frame("model"= row.names(tt), 
                     tt, "AICw"= ttt), desc(AICw))
model.summ

```


```{r gof, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# Goodness of fit.
# 
gof.m80.ergm02 <- gof(m80.ergm.fit02)
par(mfrow=c(1, 3))
plot(gof.m80.ergm02)

gof.m80.ergm1 <- gof(m80.ergm.fit1)
par(mfrow=c(1, 3))
plot(gof.m80.ergm1)

gof.m80.ergm2 <- gof(m80.ergm.fit2)
par(mfrow=c(1, 3))
plot(gof.m80.ergm2)

gof.m80.ergm3 <- gof(m80.ergm.fit3)
par(mfrow=c(1, 3))
plot(gof.m80.ergm3)

gof.m80.ergm11 <- gof(m80.ergm.fit11)
par(mfrow=c(1, 3))
plot(gof.m80.ergm11)

gof.m80.ergm12 <- gof(m80.ergm.fit12)
par(mfrow=c(1, 3))
plot(gof.m80.ergm12)

```


This additional analysis is similar to the previous, but follows:
Luke, D.A. 2015. _A User’s Guide to Network Analysis in R._ Springer, Berlin. (ISBN 978-3-319-23882-1).


```{r Testing2, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# An additional, redundant test following Luke's book.
library(ergm)
DSmod0 <- ergm(m80.s ~ edges,
    control=control.ergm(seed=40))
summary(DSmod0)

DSmod1 <- ergm(m80.s ~ edges + 
                        nodecov('tree_mass') +
                        nodecov('leaf_P') +
                        nodecov('height') +
                        nodecov('leaf_N') +
                        nodecov('NP_ratio') ,
               control=control.ergm(seed=40))
summary(DSmod1)
DSmod2 <- ergm(m80.s ~ edges + 
                        nodecov('tree_mass') +
                #       nodecov('leaf_P') +
                        nodecov('height'),
                #       nodecov('leaf_N') +
                #       nodecov('NP_ratio') ,
               control=control.ergm(seed=40))
summary(DSmod2)

anova.ergm(DSmod0, DSmod1, test="Chi")
anova.ergm(DSmod1, DSmod2, test="Chi")

```


```{r AICw_Testing2}

## Compute table
require(MuMIn)
tt<- round(BIC(DSmod0, DSmod1, DSmod2), 10)
ttt<- round(Weights(AICc(DSmod0, DSmod1, DSmod2)), 10)
data.frame(tt, "AICw"= ttt)

```



```{r Testing3, echo=TRUE, message=FALSE, warning=FALSE, fig.width=9}
# Some relationships of node parameters with predictors.
# 

tt<- data.frame("tree_mass"= m80.s %v% 'tree_mass', 
                "ttdegree"= sna::degree(m80.s, gmode='graph'),
                "tteigcentr"= sna::evcent(m80.s, rescale = F, gmode='graph'),
                "ttclos"= sna::evcent(m80.s, rescale = T, gmode='graph'))
p1<- ggplot(tt, aes(log(tree_mass), log(ttdegree))) + 
        geom_point(size=3) + geom_smooth(span = 1.5)

p2<- ggplot(tt, aes(log(tree_mass), log(tteigcentr))) + 
        geom_point(size=3) + geom_smooth(span = 1.5)

p3<- ggplot(tt, aes(log(tree_mass), log(ttclos))) + 
        geom_point(size=3) + geom_smooth(span = 1.5)

multiplot(p1, p2, p3, cols=3)

```



```{r sessioninfo, echo=TRUE}
sessionInfo()
```

