---
title: "Models"
author: "M. Florencia Miguel"
date: "25 de mayo de 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MODELS
 
 The Bernoulli or random model assumes that, for a given pair of nodes, the presence or absence of an edge between that pair is independent of the status of possible edges between any other pairs of nodes.
 
 These three statistics can be used in ergm by specifying terms altkstar, gwdegree, or gwesp, respectively, in the model. 
 
 For all these analyses I've followed:
 Kolaczyk, E.D. and  Csardi, G. 2014. _Statistical Analysis of Network Data with R_. Springer, Berlin. (ISBN 978-1-4939-0982-7). See `library(sand)`.
 
 ##### Build data
 #### Data set for models building is from ecological_attributes.csv
 
```{r initialize, message= FALSE, warning=FALSE}

library(sand)
library(ergm)
library(igraph)
library(ggplot2)

```
 
```{r}
mod<- read.csv("../data/canonical.correlations.csv",header=T,sep=",",dec=".",na.strings="NA")
str(mod)

```
```{r spatial autocorrelation, echo= TRUE, message= TRUE, warning= TRUE}
library(nlme)

#Pinheiro, J.C., and Bates, D.M. (2000) "Mixed-Effects Models in S and S-PLUS", Springer.

fm1.lme<-lme(Gra.gri ~site, random=~1|trees,data=mod)
fm1.lme

summary(fm1.lme)

fm2.lme<-update(fm1.lme, correlation=corAR1())
fm2.lme

summary(fm2.lme)

anova(fm1.lme, fm2.lme)
# Model df      AIC      BIC    logLik   Test L.Ratio p-value
#fm1.lme     1  4 2396.599 2409.370 -1194.299                       
#fm2.lme     2  5 2398.599 2414.563 -1194.299 1 vs 2       0       1

plot(fm2.lme)
plot(fm1.lme)
```

```{r interaction correlations,echo= TRUE, message= TRUE, warning= TRUE }

#http://www.statisticssolutions.com/table-of-critical-values-pearson-correlation/
#for a length of 12 consider a significant correlation higher than 0.58  (p< 0.05)

x <- mod[,60:71]

cor(x, use="complete.obs", method="kendall") 
```


#GRAZED LAND USES

```{r data.grazed.build, echo= TRUE, message= TRUE, warning= TRUE, fig.width= 9}
 
# Add the node attributes from tree traits
#vistis 31:42; fruits consumed 43:54; total effect 59:70
 
grazed<- mod %>% 
           dplyr::filter(site=="grazed") %>%  
           dplyr::select(1,60:71) 
           
           

netG.g<- graph_from_incidence_matrix(grazed, weighted= T, add.names=NULL)
nG.edg<- as_edgelist(netG.g)

 
 # Subsetting the covariates from the main dataset for each site separately.
 # 4:24 microhabitat variables, 25 to 29 mesohabitat variables
 mG_attr<-mod %>%  
           dplyr::filter(site=="grazed") %>% 
           dplyr::select(c(1,4:29))
           
 
 # Create the attributes matrix
 V(netG.g)$trees<-     mG_attr$trees
 V(netG.g)$height<-  mG_attr$height
 V(netG.g)$cr.fru<-     mG_attr$cr.fru
 V(netG.g)$gr.fru<-     mG_attr$gr.fru
 V(netG.g)$neigh<-     mG_attr$neigh
 V(netG.g)$Pros.neigh<-   mG_attr$Pros.neigh
 V(netG.g)$veg.cov.x<-   mG_attr$veg.cov.x
 V(netG.g)$veg.cov.sd<-   mG_attr$veg.cov.sd
 V(netG.g)$bare.cov.x<-   mG_attr$bare.cov.x 
 V(netG.g)$bare.cov.sd<-   mG_attr$bare.cov.sd
 V(netG.g)$litt.cov.x<-   mG_attr$litt.cov.x 
 V(netG.g)$litt.cov.sd<-   mG_attr$litt.cov.sd
 V(netG.g)$veg.cov.x1<-   mG_attr$veg.cov.x1
 V(netG.g)$veg.cov.sd1<-   mG_attr$veg.cov.sd1
 V(netG.g)$veg.cov.x2<-   mG_attr$veg.cov.x2
 V(netG.g)$veg.cov.sd2<-   mG_attr$veg.cov.sd2
 V(netG.g)$veg.cov.x3<-   mG_attr$veg.cov.x3
 V(netG.g)$veg.cov.sd3<-   mG_attr$veg.cov.sd3
 V(netG.g)$H.shannon<-   mG_attr$H.shannon 
 V(netG.g)$litt.site<-   mG_attr$litt.site 
 V(netG.g)$bare.site<-   mG_attr$bare.site 
 V(netG.g)$low.site<-   mG_attr$low.site 
 V(netG.g)$med.site<-   mG_attr$med.site 
 V(netG.g)$high.site<-   mG_attr$high.site 
 
```

##### Set matrices and graphs for model testing
 
 For now just the grazed graph is analyzed.
 
```{r Bernoulli grazed, echo=TRUE, message= TRUE, warning= TRUE, fig.width=9}
 #First separating the network into adjacency matrix and attributes
 #Loading vertex attributes for grazed
 # 
library(sand)
 AA <- get.adjacency(netG.g)
 v.attrs <- get.data.frame(netG.g, what="vertices")[1:114,]
 
 #Then, creating the analogous network object for ergm
 library(ergm) # Will load package 'network' as well.
 grazed.s <- network::as.network(as.matrix(AA),
                              directed=FALSE, bipartite=114) 
 network::set.vertex.attribute(grazed.s, "height",
                                 v.attrs$height) 
 network::set.vertex.attribute(grazed.s, "crown.fruits",
                                 v.attrs$cr.fru) 
 network::set.vertex.attribute(grazed.s, "neighbours",
                                 v.attrs$neigh)
 network::set.vertex.attribute(grazed.s, "Prosopis.neighbours",
                                 v.attrs$Pros.neigh)
 network::set.vertex.attribute(grazed.s, "vegetation.cover",
                                 v.attrs$veg.cov.x)
 network::set.vertex.attribute(grazed.s, "bare.sd",
                                 v.attrs$bare.cov.sd)
 network::set.vertex.attribute(grazed.s, "litter.sd",
                                 v.attrs$litt.cov.sd)
 network::set.vertex.attribute(grazed.s, "vegetation.sd1",
                                 v.attrs$veg.cov.sd1)
 network::set.vertex.attribute(grazed.s, "vegetation.sd2",
                                 v.attrs$veg.cov.sd2)
 network::set.vertex.attribute(grazed.s, "vegetation.sd3",
                                 v.attrs$veg.cov.sd3)
 network::set.vertex.attribute(grazed.s, "litter.site",
                                 v.attrs$litt.site)
 network::set.vertex.attribute(grazed.s, "bare.site",
                                 v.attrs$bare.site)
 network::set.vertex.attribute(grazed.s, "low.site",
                                 v.attrs$low.site)
 
 # Specifying Random Bernouilli model
 my.ergm.bern <- formula(grazed.s ~ edges)
 my.ergm.bern
 grazed.s ~ edges
 summary.statistics(my.ergm.bern) #edges 995
 
 # Models with no explanatory variables
 # 
 grazed.ergm00 <- formula(grazed.s ~ edges)                      # Bernoulli
 
 grazed.ergm01 <- formula(grazed.s ~ edges + kstar(2)            # kstar and triangle
                             + kstar(3) + triangle)
 
grazed.ergm02 <- formula(grazed.s ~ edges     # Geometrically weighted degree count
                     + gwesp(1, fixed=TRUE))
 
 summary.statistics(grazed.ergm00)#edges 1368
 summary.statistics(grazed.ergm01)#edges 1368; kstar2 84816; kstar3 2910648, triangle 0
 summary.statistics(grazed.ergm02)#edges 1368, gwesp.fixed.1 0
 
```

 
```{r models fit grazed, echo=TRUE, message= TRUE, warning= TRUE}

# Models with explanatory variables. Grazed.
#
grazed.ergm1 <- formula(grazed.s ~ edges  
                        + nodemain("height")
                        + nodemain("crown.fruits")
                        + nodemain("neighbours")
                        + nodemain("Prosopis.neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("bare.sd")
                        + nodemain("litter.sd")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.sd2")
                        + nodemain("vegetation.sd3")
                        + nodemain("litter.site")
                        + nodemain("bare.site")
                        + nodemain("low.site"))
      #+ gwesp(log(3), fixed=TRUE)                  

#explanation of transitivity https://www.sci.unich.it/~francesc/teaching/network/transitivity.html

grazed.ergm2 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height")
                        + nodemain("crown.fruits")
                        + nodemain("neighbours")
                        + nodemain("Prosopis.neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("bare.sd")
                        + nodemain("litter.sd")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.sd2")
                        + nodemain("vegetation.sd3"))
                     #   + nodemain("litter.site")
                    #    + nodemain("bare.site")
                     #   + nodemain("low.site"))

grazed.ergm3 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                   #     + nodemain("height")
                    #    + nodemain("crown.fruits")
                    #    + nodemain("neighbours")
                    #    + nodemain("Prosopis.neighbours")
                    #    + nodemain("vegetation.cover")
                    #    + nodemain("bare.sd")
                    #    + nodemain("litter.sd")
                    #    + nodemain("vegetation.sd1")
                    #    + nodemain("vegetation.sd2")
                    #    + nodemain("vegetation.sd3")
                        + nodemain("litter.site")
                        + nodemain("bare.site")
                        + nodemain("low.site"))
                    

grazed.ergm4 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height")
                        + nodemain("crown.fruits"))
                 #       + nodemain("neighbours")
                #        + nodemain("Prosopis.neighbours")
                 #       + nodemain("vegetation.cover")
                  #      + nodemain("bare.sd")
                   #     + nodemain("litter.sd")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.sd2")
                  #      + nodemain("vegetation.sd3")
                  #      + nodemain("litter.site")
                  #      + nodemain("bare.site")
                  #      + nodemain("low.site"))

grazed.ergm5 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                   #     + nodemain("height")
                  #      + nodemain("crown.fruits")
                        + nodemain("neighbours")
                        + nodemain("Prosopis.neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("bare.sd")
                        + nodemain("litter.sd")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.sd2")
                        + nodemain("vegetation.sd3"))
                   #     + nodemain("litter.site")
                  #      + nodemain("bare.site")
                  #      + nodemain("low.site"))

grazed.ergm6 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                    #    + nodemain("height")
                    #    + nodemain("crown.fruits")
                        + nodemain("neighbours")
                        + nodemain("Prosopis.neighbours"))
                    #    + nodemain("vegetation.cover")
                    #    + nodemain("bare.sd")
                    #    + nodemain("litter.sd")
                    #    + nodemain("vegetation.sd1")
                    #    + nodemain("vegetation.sd2")
                    #    + nodemain("vegetation.sd3")
                    #    + nodemain("litter.site")
                    #    + nodemain("bare.site")
                    #    + nodemain("low.site"))

grazed.ergm7 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                #        + nodemain("height")
                 #       + nodemain("crown.fruits")
                #        + nodemain("neighbours")
                #        + nodemain("Prosopis.neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("bare.sd")
                        + nodemain("litter.sd")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.sd2")
                        + nodemain("vegetation.sd3"))
                 #       + nodemain("litter.site")
                #        + nodemain("bare.site")
                #        + nodemain("low.site")) 

grazed.ergm8 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                   #     + nodemain("height")
                        + nodemain("crown.fruits"))
                  #      + nodemain("neighbours")
                  #      + nodemain("Prosopis.neighbours")
                  #      + nodemain("vegetation.cover")
                  #      + nodemain("bare.sd")
                  #      + nodemain("litter.sd")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.sd2")
                  #      + nodemain("vegetation.sd3")
                  #      + nodemain("litter.site")
                  #      + nodemain("bare.site")
                  #      + nodemain("low.site"))

grazed.ergm9 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height"))
                    #    + nodemain("crown.fruits")
                    #    + nodemain("neighbours")
                    #    + nodemain("Prosopis.neighbours")
                    #    + nodemain("vegetation.cover")
                    #    + nodemain("bare.sd")
                    #    + nodemain("litter.sd")
                    #    + nodemain("vegetation.sd1")
                    #    + nodemain("vegetation.sd2")
                    #    + nodemain("vegetation.sd3")
                    #    + nodemain("litter.site")
                    #    + nodemain("bare.site")
                    #    + nodemain("low.site"))
grazed.ergm10 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                  #      + nodemain("height")
                  #      + nodemain("crown.fruits")
                  #      + nodemain("neighbours")
                        + nodemain("Prosopis.neighbours"))
                  #      + nodemain("vegetation.cover")
                  #      + nodemain("bare.sd")
                  #      + nodemain("litter.sd")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.sd2")
                  #      + nodemain("vegetation.sd3")
                  #      + nodemain("litter.site")
                  #      + nodemain("bare.site")
                  #      + nodemain("low.site"))
                  
grazed.ergm11 <- formula(grazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
               #         + nodemain("height")
              #          + nodemain("crown.fruits")
                        + nodemain("neighbours"))
               #         + nodemain("Prosopis.neighbours")
              #          + nodemain("vegetation.cover")
              #          + nodemain("bare.sd")
              #          + nodemain("litter.sd")
              #          + nodemain("vegetation.sd1")
              #          + nodemain("vegetation.sd2")
              #          + nodemain("vegetation.sd3")
              #          + nodemain("litter.site")
              #          + nodemain("bare.site")
              #          + nodemain("low.site"))
                      
```


```{r model tests grazed, echo=TRUE, message= TRUE, warning= TRUE}
# Model tests
# 
set.seed(42)
grazed.ergm.fit00 <- ergm(grazed.ergm00)  #Bernoulli model
grazed.ergm.fit02 <- ergm(grazed.ergm02)
grazed.ergm.fit1 <- ergm(grazed.ergm1)    # Full model
grazed.ergm.fit2 <- ergm(grazed.ergm2)    #with microhabitat variables
grazed.ergm.fit3 <- ergm(grazed.ergm3)    #with mesohabitat variables 
grazed.ergm.fit4 <- ergm(grazed.ergm4)    #with intrinsic variables
grazed.ergm.fit5 <- ergm(grazed.ergm5)    #with extrinsic variables
grazed.ergm.fit6 <- ergm(grazed.ergm6)    #neighborhood variables
grazed.ergm.fit7 <- ergm(grazed.ergm7)    #vegetation covers
grazed.ergm.fit8 <- ergm(grazed.ergm8)    #fecundity variables
grazed.ergm.fit9 <- ergm(grazed.ergm9)    #size variables
grazed.ergm.fit10 <- ergm(grazed.ergm10)    #intraspecific neighborhood
grazed.ergm.fit11 <- ergm(grazed.ergm11)    #interspecific neighborhood

# NOT RUN
# mcmc.diagnostics(m80.ergm.fit0)
# mcmc.diagnostics(m80.ergm.fit1)
# mcmc.diagnostics(m80.ergm.fit2)
# mcmc.diagnostics(m80.ergm.fit3)
# mcmc.diagnostics(m80.ergm.fit11)
# mcmc.diagnostics(m80.ergm.fit12)


# in order to know the effect of the covariates used at each model, I performed anova for each model
anova.ergm(grazed.ergm.fit00, test="Chi") #page 92 book
anova.ergm(grazed.ergm.fit02, test="Chi")
anova.ergm(grazed.ergm.fit1, test="Chi")
anova.ergm(grazed.ergm.fit2, test="Chi")
anova.ergm(grazed.ergm.fit3, test="Chi")
anova.ergm(grazed.ergm.fit4, test="Chi")
anova.ergm(grazed.ergm.fit5, test="Chi")
anova.ergm(grazed.ergm.fit6, test="Chi")
anova.ergm(grazed.ergm.fit7, test="Chi")
anova.ergm(grazed.ergm.fit8, test="Chi")
anova.ergm(grazed.ergm.fit9, test="Chi")
anova.ergm(grazed.ergm.fit10, test="Chi")
anova.ergm(grazed.ergm.fit11, test="Chi")


anova.ergm(grazed.ergm.fit00, grazed.ergm.fit1, test="Chi")
anova.ergm(grazed.ergm.fit02, grazed.ergm.fit1, test="Chi")
anova.ergm(grazed.ergm.fit1, grazed.ergm.fit2, test="Chi")
anova.ergm(grazed.ergm.fit1, grazed.ergm.fit3, test="Chi")
anova.ergm(grazed.ergm.fit2, grazed.ergm.fit3, test="Chi")
anova.ergm(grazed.ergm.fit2, grazed.ergm.fit4, test="Chi")
anova.ergm(grazed.ergm.fit2, grazed.ergm.fit5, test="Chi")
anova.ergm(grazed.ergm.fit4, grazed.ergm.fit5, test="Chi")
anova.ergm(grazed.ergm.fit4, grazed.ergm.fit7, test="Chi")
anova.ergm(grazed.ergm.fit4, grazed.ergm.fit8, test="Chi")
anova.ergm(grazed.ergm.fit4, grazed.ergm.fit9, test="Chi")
anova.ergm(grazed.ergm.fit5, grazed.ergm.fit6, test="Chi")
anova.ergm(grazed.ergm.fit5, grazed.ergm.fit7, test="Chi")
anova.ergm(grazed.ergm.fit6, grazed.ergm.fit7, test="Chi")
anova.ergm(grazed.ergm.fit7, grazed.ergm.fit3, test="Chi")
anova.ergm(grazed.ergm.fit8, grazed.ergm.fit9, test="Chi")
anova.ergm(grazed.ergm.fit10, grazed.ergm.fit11, test="Chi")



summary.ergm(grazed.ergm.fit00) 
summary.ergm(grazed.ergm.fit02)
summary.ergm(grazed.ergm.fit1) #interpretation of estimates page 94 book
summary.ergm(grazed.ergm.fit2)
summary.ergm(grazed.ergm.fit3)
summary.ergm(grazed.ergm.fit4)
summary.ergm(grazed.ergm.fit5)
summary.ergm(grazed.ergm.fit6)
summary.ergm(grazed.ergm.fit7)
summary.ergm(grazed.ergm.fit8)
summary.ergm(grazed.ergm.fit9)
summary.ergm(grazed.ergm.fit10)
summary.ergm(grazed.ergm.fit11)


```


```{r}
## Compute table
require(MuMIn)
library(dplyr)

tt<- round(BIC(grazed.ergm.fit00, grazed.ergm.fit02, grazed.ergm.fit1, grazed.ergm.fit2,
               grazed.ergm.fit3, grazed.ergm.fit4, grazed.ergm.fit5,grazed.ergm.fit6, grazed.ergm.fit7,            grazed.ergm.fit9, grazed.ergm.fit10, grazed.ergm.fit11), 10)

ttt<- round(Weights(AICc(grazed.ergm.fit00, grazed.ergm.fit02, grazed.ergm.fit1, grazed.ergm.fit2,
               grazed.ergm.fit3, grazed.ergm.fit4, grazed.ergm.fit5, grazed.ergm.fit6, grazed.ergm.fit7,            grazed.ergm.fit9, grazed.ergm.fit10, grazed.ergm.fit11)), 10)

model.summ<- arrange(data.frame("model"= row.names(tt), 
                     tt, "AICw"= ttt), desc(AICw))
model.summ

```

```{r}
# Goodness of fit.
# 
gof.grazed.ergm02 <- gof(grazed.ergm.fit02)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm02)

gof.grazed.ergm1 <- gof(grazed.ergm.fit1)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm1)

gof.grazed.ergm2 <- gof(grazed.ergm.fit2)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm2)

gof.grazed.ergm3 <- gof(grazed.ergm.fit3)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm3)

gof.grazed.ergm4 <- gof(grazed.ergm.fit4)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm4)

gof.grazed.ergm5 <- gof(grazed.ergm.fit5)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm5)

gof.grazed.ergm6 <- gof(grazed.ergm.fit6)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm6)

gof.grazed.ergm7 <- gof(grazed.ergm.fit7)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm7)

gof.grazed.ergm8 <- gof(grazed.ergm.fit8)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm8)

gof.grazed.ergm9 <- gof(grazed.ergm.fit9)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm9)

gof.grazed.ergm10 <- gof(grazed.ergm.fit10)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm10)

gof.grazed.ergm11 <- gof(grazed.ergm.fit11)
par(mfrow=c(1, 3))
plot(gof.grazed.ergm11)

```


```{r}
# Some relationships of node parameters with predictors.
# 

tt<- data.frame("height"= grazed.s %v% 'height', 
                "ttdegree"= sna::degree(grazed.s, gmode='graph'),
                "tteigcentr"= sna::evcent(grazed.s, rescale = F, gmode='graph'),
                "ttclos"= sna::evcent(grazed.s, rescale = T, gmode='graph'))
p1<- ggplot(tt, aes(log(height), log(ttdegree))) + 
        geom_point(size=3) + geom_smooth(span = 1.5)

p2<- ggplot(tt, aes(log('Vegetation cover'), log(tteigcentr))) + 
        geom_point(size=3) + geom_smooth(span = 1.5)

p3<- ggplot(tt, aes(log('Vegetation cover'), log(ttclos))) + 
        geom_point(size=3) + geom_smooth(span = 1.5)

p1
p2
p3

multiplot(p1, p2, p3, cols=3)

```

#UNGRAZED LAND USES

```{r data.ungrazed.build, echo= TRUE, message= TRUE, warning= TRUE, fig.width= 9}
 
# Add the node attributes from tree traits
 # 
 
ungrazed<- mod %>% 
           dplyr::filter(site=="ungrazed") %>%  
           dplyr::select(1,60:71) 
           
           

netUG.g<- graph_from_incidence_matrix(ungrazed, weighted= T, add.names=NULL)
nUG.edg<- as_edgelist(netUG.g)

 
 # Subsetting the covariates from the main dataset for each site separately.
 # 
 mUG_attr<-mod %>%  
           dplyr::filter(site=="ungrazed") %>% 
           dplyr::select(c(1,4:29))
           
 
 # Create the attributes matrix
 V(netUG.g)$trees<-     mUG_attr$trees
 V(netUG.g)$height<-  mUG_attr$height
 V(netUG.g)$cr.fru<-     mUG_attr$cr.fru
 V(netUG.g)$gr.fru<-     mUG_attr$gr.fru
 V(netUG.g)$neigh<-     mUG_attr$neigh
 V(netUG.g)$Pros.neigh<-   mUG_attr$Pros.neigh
 V(netUG.g)$veg.cov.x<-   mUG_attr$veg.cov.x
 V(netUG.g)$veg.cov.sd<-   mUG_attr$veg.cov.sd
 V(netUG.g)$bare.cov.x<-   mUG_attr$bare.cov.x 
 V(netUG.g)$bare.cov.sd<-   mUG_attr$bare.cov.sd
 V(netUG.g)$litt.cov.x<-   mUG_attr$litt.cov.x 
 V(netUG.g)$litt.cov.sd<-   mUG_attr$litt.cov.sd
 V(netUG.g)$veg.cov.x1<-   mUG_attr$veg.cov.x1
 V(netUG.g)$veg.cov.sd1<-   mUG_attr$veg.cov.sd1
 V(netUG.g)$veg.cov.x2<-   mUG_attr$veg.cov.x2
 V(netUG.g)$veg.cov.sd2<-   mUG_attr$veg.cov.sd2
 V(netUG.g)$veg.cov.x3<-   mUG_attr$veg.cov.x3
 V(netUG.g)$veg.cov.sd3<-   mUG_attr$veg.cov.sd3
 V(netUG.g)$H.shannon<-   mUG_attr$H.shannon 
 V(netUG.g)$litt.site<-   mUG_attr$litt.site 
 V(netUG.g)$bare.site<-   mUG_attr$bare.site 
 V(netUG.g)$low.site<-   mUG_attr$low.site 
 V(netUG.g)$med.site<-   mUG_attr$med.site 
 V(netUG.g)$high.site<-   mUG_attr$high.site

```

##### Set matrices and graphs for model testing
 
 For now just the grazed graph is analyzed.
 
```{r Bernoulli ungrazed, echo=TRUE, message= TRUE, warning= TRUE, fig.width=9}
 #First separating the network into adjacency matrix and attributes
 #Loading vertex attributes for grazed
 # 
library(sand)
 AA <- get.adjacency(netUG.g)
 v.attrs <- get.data.frame(netUG.g, what="vertices")[1:68,]
 
 #Then, creating the analogous network object for ergm
 library(ergm) # Will load package 'network' as well.
 ungrazed.s <- network::as.network(as.matrix(AA),
                              directed=FALSE, bipartite=68) 
 network::set.vertex.attribute(ungrazed.s, "height",
                                 v.attrs$height) 
 network::set.vertex.attribute(ungrazed.s, "crown.fruits",
                                 v.attrs$cr.fru) 
 network::set.vertex.attribute(ungrazed.s, "ground.fruits",
                                 v.attrs$gr.fru)
 network::set.vertex.attribute(ungrazed.s, "neighbours",
                                 v.attrs$neigh)
 network::set.vertex.attribute(ungrazed.s, "vegetation.cover",
                                 v.attrs$veg.cov.x)
 network::set.vertex.attribute(ungrazed.s, "vegetation.sd",
                                 v.attrs$veg.cov.sd)
 network::set.vertex.attribute(ungrazed.s, "bare.cover",
                                 v.attrs$bare.cov.x)
 network::set.vertex.attribute(ungrazed.s, "vegetation.sd1",
                                 v.attrs$veg.cov.sd1)
 network::set.vertex.attribute(ungrazed.s, "vegetation.x2",
                                 v.attrs$veg.cov.x2)
 network::set.vertex.attribute(ungrazed.s, "litter.site",
                                 v.attrs$litt.site)
 network::set.vertex.attribute(ungrazed.s, "medium.site",
                                 v.attrs$med.site)
 # Specifying Random Bernouilli model
 my.ergm.bern <- formula(ungrazed.s ~ edges)
 my.ergm.bern
 ungrazed.s ~ edges
 summary.statistics(my.ergm.bern) #edges 995
 
 # Models with no explanatory variables
 # 
 ungrazed.ergm00 <- formula(ungrazed.s ~ edges)                      # Bernoulli
 
 ungrazed.ergm01 <- formula(ungrazed.s ~ edges + kstar(2)            # kstar and triangle
                             + kstar(3) + triangle)
 
ungrazed.ergm02 <- formula(ungrazed.s ~ edges     # Geometrically weighted degree count
                     + gwesp(1, fixed=TRUE))
 
 summary.statistics(ungrazed.ergm00)#edges 482
 summary.statistics(ungrazed.ergm01)#edges 482; kstar2 17420; kstar3 353319, triangle 0
 summary.statistics(ungrazed.ergm02)#edges 482, gwesp.fixed.1 0
 
```

 
```{r models fit ungrazed, echo=TRUE, message= TRUE, warning= TRUE}

# Models with explanatory variables. Grazed.
#
ungrazed.ergm1 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height")
                        + nodemain("crown.fruits")
                        + nodemain("ground.fruits")
                        + nodemain("neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("vegetation.sd")
                        + nodemain("bare.cover")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.x2")
                        + nodemain("litter.site")
                        + nodemain("medium.site"))

#explanation of transitivity https://www.sci.unich.it/~francesc/teaching/network/transitivity.html

ungrazed.ergm2 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height")
                        + nodemain("crown.fruits")
                        + nodemain("ground.fruits")
                        + nodemain("neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("vegetation.sd")
                        + nodemain("bare.cover")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.x2"))
                   #     + nodemain("litter.site")
                  #      + nodemain("medium.site"))


ungrazed.ergm3 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                     #   + nodemain("height")
                    #    + nodemain("crown.fruits")
                    #    + nodemain("ground.fruits")
                    #    + nodemain("neighbours")
                    #    + nodemain("vegetation.cover")
                    #    + nodemain("vegetation.sd")
                    #    + nodemain("bare.cover")
                    #    + nodemain("vegetation.sd1")
                    #    + nodemain("vegetation.x2")
                        + nodemain("litter.site")
                        + nodemain("medium.site"))

 

ungrazed.ergm4 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height")
                        + nodemain("crown.fruits")
                        + nodemain("ground.fruits"))
                   #     + nodemain("neighbours")
                  #      + nodemain("vegetation.cover")
                  #      + nodemain("vegetation.sd")
                  #      + nodemain("bare.cover")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.x2")
                  #      + nodemain("litter.site")
                  #      + nodemain("medium.site"))



ungrazed.ergm5 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                    #    + nodemain("height")
                    #    + nodemain("crown.fruits")
                    #    + nodemain("ground.fruits")
                        + nodemain("neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("vegetation.sd")
                        + nodemain("bare.cover")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.x2"))
                    #    + nodemain("litter.site")
                    #    + nodemain("medium.site"))


ungrazed.ergm6 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                  #      + nodemain("height")
                  #      + nodemain("crown.fruits")
                  #      + nodemain("ground.fruits")
                        + nodemain("neighbours"))
                  #      + nodemain("vegetation.cover")
                  #      + nodemain("vegetation.sd")
                  #      + nodemain("bare.cover")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.x2")
                  #      + nodemain("litter.site")
                  #      + nodemain("med.site"))

ungrazed.ergm7 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                #        + nodemain("height")
                #        + nodemain("crown.fruits")
                #        + nodemain("ground.fruits")
                #        + nodemain("neighbours")
                        + nodemain("vegetation.cover")
                        + nodemain("vegetation.sd")
                        + nodemain("bare.cover")
                        + nodemain("vegetation.sd1")
                        + nodemain("vegetation.x2"))
                #        + nodemain("litter.site")
                #        + nodemain("medium.site"))


ungrazed.ergm8 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                   #     + nodemain("height")
                        + nodemain("crown.fruits")
                        + nodemain("ground.fruits"))
                  #      + nodemain("neighbours")
                  #      + nodemain("vegetation.cover")
                  #      + nodemain("vegetation.sd")
                  #      + nodemain("bare.cover")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.x2")
                  #      + nodemain("litter.site")
                  #      + nodemain("medium.site"))

ungrazed.ergm9 <- formula(ungrazed.s ~ edges + gwesp(log(3), fixed=TRUE) 
                        + nodemain("height"))
                   #     + nodemain("crown.fruits")
                  #      + nodemain("ground.fruits")
                  #      + nodemain("neighbours")
                  #      + nodemain("vegetation.cover")
                  #      + nodemain("vegetation.sd")
                  #      + nodemain("bare.cover")
                  #      + nodemain("vegetation.sd1")
                  #      + nodemain("vegetation.x2")
                  #      + nodemain("litter.site")
                  #      + nodemain("medium.site"))



```


```{r model tests ungrazed, echo=TRUE, message= TRUE, warning= TRUE}
# Model tests
# 
set.seed(42)
ungrazed.ergm.fit00 <- ergm(ungrazed.ergm00)
ungrazed.ergm.fit02 <- ergm(ungrazed.ergm02)
ungrazed.ergm.fit1 <- ergm(ungrazed.ergm1)    # Full model
ungrazed.ergm.fit2 <- ergm(ungrazed.ergm2)    #with microhabitat variables
ungrazed.ergm.fit3 <- ergm(ungrazed.ergm3)    #with mesohabitat variables 
ungrazed.ergm.fit4 <- ergm(ungrazed.ergm4)    # intrinsic variables
ungrazed.ergm.fit5 <- ergm(ungrazed.ergm5)    #extrinsic variables
ungrazed.ergm.fit6 <- ergm(ungrazed.ergm6)    #neighborhood  
ungrazed.ergm.fit7 <- ergm(ungrazed.ergm7)    #covers
ungrazed.ergm.fit8 <- ergm(ungrazed.ergm8)    #fecundity
ungrazed.ergm.fit9 <- ergm(ungrazed.ergm9)    #tree size


# NOT RUN
# mcmc.diagnostics(m80.ergm.fit0)
# mcmc.diagnostics(m80.ergm.fit1)
# mcmc.diagnostics(m80.ergm.fit2)
# mcmc.diagnostics(m80.ergm.fit3)
# mcmc.diagnostics(m80.ergm.fit11)
# mcmc.diagnostics(m80.ergm.fit12)

anova.ergm(ungrazed.ergm.fit00, test="Chi") #page 92 book
anova.ergm(ungrazed.ergm.fit02, test="Chi")
anova.ergm(ungrazed.ergm.fit1, test="Chi")
anova.ergm(ungrazed.ergm.fit2, test="Chi")
anova.ergm(ungrazed.ergm.fit3, test="Chi")
anova.ergm(ungrazed.ergm.fit4, test="Chi")
anova.ergm(ungrazed.ergm.fit5, test="Chi")
anova.ergm(ungrazed.ergm.fit6, test="Chi")
anova.ergm(ungrazed.ergm.fit7, test="Chi")
anova.ergm(ungrazed.ergm.fit8, test="Chi")
anova.ergm(ungrazed.ergm.fit9, test="Chi")

anova.ergm(ungrazed.ergm.fit00, ungrazed.ergm.fit1, test="Chi")
anova.ergm(ungrazed.ergm.fit02, ungrazed.ergm.fit1, test="Chi")
anova.ergm(ungrazed.ergm.fit1, ungrazed.ergm.fit2, test="Chi")
anova.ergm(ungrazed.ergm.fit1, ungrazed.ergm.fit3, test="Chi")
anova.ergm(ungrazed.ergm.fit2, ungrazed.ergm.fit3, test="Chi")
anova.ergm(ungrazed.ergm.fit2, ungrazed.ergm.fit4, test="Chi")
anova.ergm(ungrazed.ergm.fit2, ungrazed.ergm.fit5, test="Chi")
anova.ergm(ungrazed.ergm.fit4, ungrazed.ergm.fit5, test="Chi")
anova.ergm(ungrazed.ergm.fit4, ungrazed.ergm.fit7, test="Chi")
anova.ergm(ungrazed.ergm.fit4, ungrazed.ergm.fit8, test="Chi")
anova.ergm(ungrazed.ergm.fit4, ungrazed.ergm.fit9, test="Chi")
anova.ergm(ungrazed.ergm.fit5, ungrazed.ergm.fit6, test="Chi")
anova.ergm(ungrazed.ergm.fit5, ungrazed.ergm.fit7, test="Chi")
anova.ergm(ungrazed.ergm.fit6, ungrazed.ergm.fit7, test="Chi")
anova.ergm(ungrazed.ergm.fit7, ungrazed.ergm.fit3, test="Chi")
anova.ergm(ungrazed.ergm.fit8, ungrazed.ergm.fit9, test="Chi")


summary.ergm(ungrazed.ergm.fit00) 
summary.ergm(ungrazed.ergm.fit02)
summary.ergm(ungrazed.ergm.fit1) #interpretation of estimates page 94 book
summary.ergm(ungrazed.ergm.fit2)
summary.ergm(ungrazed.ergm.fit3)
summary.ergm(ungrazed.ergm.fit4)
summary.ergm(ungrazed.ergm.fit5)
summary.ergm(ungrazed.ergm.fit6)
summary.ergm(ungrazed.ergm.fit7)
summary.ergm(ungrazed.ergm.fit8)
summary.ergm(ungrazed.ergm.fit9)

```

```{r}
## Compute table
require(MuMIn)
library(dplyr)

tt<- round(BIC(ungrazed.ergm.fit00, ungrazed.ergm.fit02, ungrazed.ergm.fit1, ungrazed.ergm.fit2,
               ungrazed.ergm.fit3, ungrazed.ergm.fit4, ungrazed.ergm.fit5,ungrazed.ergm.fit6, ungrazed.ergm.fit7, ungrazed.ergm.fit8, ungrazed.ergm.fit9), 10)

ttt<- round(Weights(AICc(ungrazed.ergm.fit00, ungrazed.ergm.fit02, ungrazed.ergm.fit1, ungrazed.ergm.fit2,
               ungrazed.ergm.fit3, ungrazed.ergm.fit4, ungrazed.ergm.fit5,ungrazed.ergm.fit6, ungrazed.ergm.fit7, ungrazed.ergm.fit8, ungrazed.ergm.fit9)), 10)

model.summ<- arrange(data.frame("model"= row.names(tt), 
                     tt, "AICw"= ttt), desc(AICw))
model.summ

```

```{r}
# Goodness of fit.
# 
gof.ungrazed.ergm02 <- gof(ungrazed.ergm.fit02)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm02)

gof.ungrazed.ergm1 <- gof(ungrazed.ergm.fit1)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm1)

gof.ungrazed.ergm2 <- gof(ungrazed.ergm.fit2)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm2)

gof.ungrazed.ergm3 <- gof(ungrazed.ergm.fit3)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm3)

gof.ungrazed.ergm4 <- gof(ungrazed.ergm.fit4)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm4)

gof.ungrazed.ergm5 <- gof(ungrazed.ergm.fit5)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm5)

gof.ungrazed.ergm6 <- gof(ungrazed.ergm.fit6)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm6)

gof.ungrazed.ergm7 <- gof(ungrazed.ergm.fit7)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm7)

gof.ungrazed.ergm8 <- gof(ungrazed.ergm.fit8)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm8)

gof.ungrazed.ergm9 <- gof(ungrazed.ergm.fit9)
par(mfrow=c(1, 3))
plot(gof.ungrazed.ergm9)


```
