---
title: "Models"
author: "M. Florencia Miguel"
date: "25 de mayo de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MODELS
 
 The Bernoulli or random model assumes that, for a given pair of nodes, the presence or absence of an edge between that pair is independent of the status of possible edges between any other pairs of nodes.
 
 These three statistics can be used in ergm by specifying terms altkstar, gwdegree, or gwesp, respectively, in the model. 
 
 For all these analyses I've followed:
 Kolaczyk, E.D. and  Cs?rdi, G. 2014. _Statistical Analysis of Network Data with R_. Springer, Berlin. (ISBN 978-1-4939-0982-7). See `library(sand)`.
 
 ##### Build data
 #### Data set for models building is from ecological_attributes.csv 
 
```{r}
traits<- read.csv("../data/ecological_attributes.csv",header=T,sep=",",dec=".",na.strings="NA")
str(traits)
```
 
```{r data.build, echo= TRUE, message= TRUE, warning= TRUE, fig.width= 9}
 
# Add the node attributes
 # 
 netwG.g<- graph_from_incidence_matrix(grazed, weighted= T, add.names=NULL)
 nG.edg<- as_edgelist(netwG.g)
 
 # Subsetting the covariates from the main dataset for each site separately.
 # 
 mG_attr<-main %>%  
           dplyr::filter(site=="grazed") %>% 
           dplyr::select(c(1,4:23)) 
 
 # Create the attributes matrix
 V(netwG.g)$trees<-     mG_attr$tid
 V(netwG.g)$height<-  mG_attr$height
 V(netwG.g)$cr.fru<-     mG_attr$cf
 V(netwG.g)$gr.fru<-     mG_attr$gf
 V(netwG.g)$neigh<-     mG_attr$neigh
 V(netwG.g)$Pros.neigh<-   mG_attr$Pn
 V(netwG.g)$veg.cov.x<-   mG_attr$mvc
 V(netwG.g)$veg.cov.sd<-   mG_attr$sdvc
 V(netwG.g)$bare.cov.x<-   mG_attr$mbc  
 V(netwG.g)$bare.cov.sd<-   mG_attr$sdbc
 V(netwG.g)$litt.cov.x<-   mG_attr$mlc  
 V(netwG.g)$litt.cov.sd<-   mG_attr$sdlc
 V(netwG.g)$H.shannon<-   mG_attr$H 
 V(netwG.g)$x<-   mG_attr$x 
 V(netwG.g)$y<-   mG_attr$y 
```
##### Correlations among covariates
 
```{r correlations, echo=TRUE, message= TRUE, warning= TRUE, fig.width=8}
 # Correlations among tree characteristics
 # 
 mat<-mG_attr[,c(2:23)]
 mm<- stats::cor(mat, use="pairwise.complete.obs")
 
 library(GGally)
 p1<- ggcorr(data = NULL, cor_matrix = cor(mm[, -1], use = "everything"))
 
 p2<- ggcorr(mG_attr[,c(2:6)], geom = "blank", label = TRUE, hjust = 0.75) +
         geom_point(size = 10, aes(color = coefficient > 0, 
                    alpha = abs(coefficient) > 0.5)) +
                    scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
                    guides(color = FALSE, alpha = FALSE)
 multiplot (p1, p2, cols=2)
 
```
##### Set matrices and graphs for model testing
 
 For now just the grazed graph is analyzed.
 
```{r Bernouilli, echo=TRUE, message= TRUE, warning= TRUE, fig.width=9}
 # Loading vertex attributes for grazed
 # 
library(sand)
 AA <- get.adjacency(netwG.g)
 v.attrs <- get.data.frame(netwG.g, what="vertices")[1:42,]
 
 library(ergm) # Will load package 'network' as well.
 grazed.s <- network::as.network(as.matrix(AA),
                              directed=FALSE, bipartite=42) 
 network::set.vertex.attribute(grazed.s, "height",
                                 v.attrs$height) 
 network::set.vertex.attribute(grazed.s, "crown.fruits",
                                 v.attrs$crown.fruits) 
 network::set.vertex.attribute(grazed.s, "ground.fruits",
                                 v.attrs$ground.fruits)
 network::set.vertex.attribute(grazed.s, "neighbours",
                                 v.attrs$neighbours)
 network::set.vertex.attribute(grazed.s, "Prosopis.neighbours",
                                 v.attrs$Prosopis.neighbours)
 
 # Random Bernouilli model
 my.ergm.bern <- formula(grazed.s ~ edges)
 my.ergm.bern
 grazed.s ~ edges
 summary.statistics(my.ergm.bern)
 
 # Models with no explanatory variables
 # 
 grazed.ergm00 <- formula(grazed.s ~ edges)                      # Bernoulli
 
 grazed.ergm01 <- formula(grazed.s ~ edges + kstar(2)            # kstar
                             + kstar(3) + triangle)
 
grazed.ergm02 <- formula(grazed.s ~ edges     # Geometrically weighted degree count
                     + gwesp(1, fixed=TRUE))
 
 summary.statistics(grazed.ergm00)
 summary.statistics(grazed.ergm01)
 summary.statistics(grazed.ergm02)
 
```
 
 
 